On $K_2$-analogue of Serre problem for $\SL_4$

### 2018-09-27 paper.tex
Introduced relative Steinberg symbol

\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\UU}{\hat{\mathrm{U}}}      % Unipotent subgroup on the Steinberg level
\newcommand{\StB}{\hat{\mathrm{B}}}     % Borel subgroup on the Steinberg level
\newcommand{\StH}{\hat{\mathrm{H}}}     % Torus on the Steinberg level
\newcommand{\XX}{\mathcal{X}}           % (subset of) Steinberg generators
\newcommand{\RR}[1]{\mathcal{R}_{#1}}   % Steinberg relations
\newcommand{\ext}[1]{\mu_{#1}}           % Map from St(Ф, R, I) --> St(Ф, R)

@@

It is clear the above construction is functorial in $R$, for a ring map $f\colon R\to S$ we denote the correponding map of Steinberg groups by $f_*$.

\subsection{Some distinguished subgroups}
%TODO: definition of h_\alpha(\xi)
%TODO: definition of Steinberg symbols

We denote by $\StH(\Phi, R)$ the subgroup of $\St(\Phi, R)$ generated by all elements $h_\alpha(\xi)$, $\xi\in R^*$ and by $\StB(\Phi, R)$ the product of the two subgroups $\StH(\Phi, R)$ and $\UU(\Phi^+, R)$ inside $\St(\Phi, R)$.

By definition, $\Sym(\Phi, R)$ is the subgroup of $\St(\Phi, R)$ generated by all Steinberg symbols $\{a, b\}_\alpha$, $a, b \in R^*$. 
$\Sym(\Phi, R)$ is contained in $\K_2(\Phi, R)$ and is a central subgroup of $\St(\Phi, R)$ provided $\rk(\Phi)\geq 2$.

There is the following chain of inclusions:
\begin{equation}
  \Sym(\Phi, R) \subseteq \StH(\Phi, R) \subseteq \StB(\Phi, R) \subseteq \St(\Phi, R)
\end{equation}

\subsection{Relative Steinberg groups}
First of all, let us recall the definition of the category of pairs. 
As suggested by the notation, the objects of $\catname{Pairs}$ are pairs $(R, I)$ consisting of a unital commutative ring $R$ and an ideal $I \trianglelefteq R$.
The set of morphisms between $(R, I)$ and $(S, J)$ consists, by definition, of all ring-theoretic maps $f\colon R\to S$ such that $f(I)\subseteq J$.
There is an obvious fully faithful embedding $\iota\colon \catname{CRings} \to \catname{Pairs}$ defined on objects by $\iota(R) = (R, R)$.
%Now if there is a functor $F\colon \catname{CRings}\to \catname{Grp}$ then its {\it relativization} is another functor $\widetilde{F}\colon \catname{Pairs} \to \catname{Groups}$ such that $\widetilde{F} \iota = F$.

@@

\begin{equation}
 \xymatrix{ 1 \ar[r] & C(\Phi, R, I) \ar[r] & \St(\Phi, R, I) \ar[r]^{\ext{R}} & \St(\Phi, R) \ar[r]^{\pi_*} & \St(\Phi, R/I) \ar[r] & 1 }
\end{equation}

\begin{multline}
 \{ t, u \}'_\alpha = h_{\alpha}(u) ^ {h_{?}(t)} =  y_\alpha(t(u-1)) \cdot y_{-\alpha}\left(-t^{-1}(u^{-1}-1)\right)^{x_\alpha(-t)} \cdot y_\alpha\left(t(u-1)\right)^{w_\alpha(-t)} \cdot \\
\cdot y_\alpha(u-1) \cdot y_{-\alpha}\left(-(u^{-1}-1)\right)^{x_\alpha(-1)} \cdot y_\alpha(u-1)^{w_\alpha(-1)}.
\end{multline}

@@

\begin{rem} Despite the fact that formula~\eqref{eq:sigmadef} determines the maps $\sigma(\varpi)$, $\sigma(-\varpi)$ uniquely, 
 we cannot a priori be sure that~\eqref{eq:sigmadef} actually {\it defines} them, i.\,e. gives well-defined maps between $N(\pm\varpi)$.
This is caused by the fact there is no direct description of the set of relations which hold in $N(\pm\varpi)$ between the generators $\XX(\pm\varpi)$, so we cannot verify that $\sigma(\pm\varpi)$ preserves these relations. \end{rem}

@@

Denote by $V_0$ the set $\St(\Phi, A[t]) \times \StH(A[t, t^{-1}]) \times \St(\Phi, A[t, t^{-1}], m[t, t^{-1}])$. 
We define an equivalence relation $\sim$ on $V_0$ as the minimal equivalence relation generated by the following 3 elementary transformations:
\begin{enumerate}
 \item $(a \cdot u, b, \beta) \sim (a, j_+^*(u) \cdot b, \beta)$, where $u$ is an element $\UU(\Phi^+, A[t])$;
 \item $(a \cdot \ext{A[t]}(\gamma), b, \beta) \sim (a, b, j^*_+(\gamma)^b \cdot \beta)$, where $\gamma$ is an element of $\St(\Phi, A[t], m[t])$;
  %and $b$ acts on $j^*_+(\gamma)$ via the conjugation action of $\St(\Phi, R)$ on $\St(\Phi, A[t, t^{-1}], m[t, t^{-1}])$;
 \item $(a, b \cdot u, \beta) \sim (a, b, u\cdot \beta)$, where $u$ is an element of $\UU(\Phi^+, m[t, t^{-1}])$;
 \item $(a, b \cdot \{ t, a \}_\alpha, \beta) \sim (a, b, \{ t, a \}'_\alpha \cdot \beta) $, where $a \in (1+m)^*$
  and $\{t, a\}'_\alpha$ is a relative Steinberg symbol discussed in ???
\end{enumerate}

>>>>

\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\UU}{\hat{\mathrm{U}}}      % Unipotent subgroup on the Steinberg level
\newcommand{\StB}{\hat{\mathrm{B}}}     % Borel subgroup on the Steinberg level
\newcommand{\StH}{\hat{\mathrm{H}}}     % Torus on the Steinberg level
\newcommand{\StW}{\hat{\mathrm{W}}}     % weyl on the Steinberg level
\newcommand{\XX}{\mathcal{X}}           % (subset of) Steinberg generators
\newcommand{\RR}[1]{\mathcal{R}_{#1}}   % Steinberg relations
\newcommand{\ext}[1]{\mu_{#1}}           % Map from St(Ф, R, I) --> St(Ф, R)

@@

It is clear the above construction is functorial in $R$, for a ring map $f\colon R\to S$ we denote the corresponding map of Steinberg groups by $f_*$.

Recall that for $\alpha\in\Phi$, $\varepsilon\in R^*$ the semisimple root elements $h_\alpha(\varepsilon)$ are defined as $h_\alpha(\varepsilon)=w_\alpha(\varepsilon)w_\alpha(-1)$.
Denote by $\StW(\Phi, R)$ the subgroup of $\St(\Phi, R)$ generated by all elements
$w_\alpha(\varepsilon)$, $\varepsilon\in R^*$, $\alpha\in\Phi$, and by $\StH(\Phi,R)$
 the subgroup generated by all elements $h_\alpha(\varepsilon)$, $\varepsilon\in R^*$, $\alpha\in\Phi$.
We also denote by $\StB(\Phi, R)$ the product of the two subgroups $\StH(\Phi, R)$ and $\UU(\Phi^+, R)$ inside $\St(\Phi, R)$.

For $u, v \in R^*$ and $\alpha\in\Phi$ the element $\{u,v\}_\alpha=h_\alpha(uv)h_\alpha(u)^{-1}h_\alpha(v)^{-1}$ will be
 called a {\it Steinberg symbol}. By definition, $\Sym(\Phi, R)$ is the subgroup of $\St(\Phi, R)$ generated by all Steinberg symbols $\{u, v\}_\alpha$, $u, v \in R^*$. 
$\Sym(\Phi, R)$ is contained in $\K_2(\Phi, R)$ and is a central subgroup of $\St(\Phi, R)$ provided $\rk(\Phi)\geq 2$.

There is the following chain of obvious inclusions between the subgroups introduced above:
\begin{equation}
  \Sym(\Phi, R) \subseteq \StH(\Phi, R) \subseteq \StB(\Phi, R) \subseteq \St(\Phi, R)
\end{equation}

\subsection{Relative Steinberg groups} \label{sec:relative-steinberg}
First of all, let us recall the definition of the category of pairs. 
As suggested by the notation, the objects of $\catname{Pairs}$ are pairs $(R, I)$ consisting of a unital commutative ring $R$ and an ideal $I \trianglelefteq R$.
The set of morphisms between $(R, I)$ and $(S, J)$ consists, by definition, of all ring-theoretic maps $f\colon R\to S$ such that $f(I)\subseteq J$.
There is an obvious fully faithful embedding $\iota\colon \catname{CRings} \to \catname{Pairs}$ defined on objects by $\iota(R) = (R, R)$.
%Now if there is a functor $F\colon \catname{CRings}\to \catname{Grp}$ then its {\it relativization} is another functor $\widetilde{F}\colon \catname{Pairs} \to \catname{Groups}$ such that $\widetilde{F} \iota = F$.

@@

\begin{equation}
 \xymatrix{ 1 \ar[r] & C(\Phi, R, I) \ar[r] & \St(\Phi, R, I) \ar[r]^{\ext{I}} & \St(\Phi, R) \ar[r]^{\pi_*} & \St(\Phi, R/I) \ar[r] & 1 }
\end{equation}

%Now suppose that $t, u \in R^*$ are invertible elements such that $u-1 \in I$. 
In the sequel we will also need a {\it relative Steinberg symbol}, i.\,e. a map
\begin{equation*} \{-, -\}'_\alpha \colon R^* \times (1 + I)^* \to \St(\Phi, R, I) \end{equation*}
such that $ \ext{I} \{u, v\}'_\alpha = \{u, v\}_\alpha$. It is not hard to verify that the image under $\mu_I$ of the element
\begin{equation} f(\alpha, t, u)= y_\alpha(t(u-1)) \cdot y_{-\alpha}(-t^{-1}(u^{-1}-1))^{x_\alpha(-t)} \cdot y_{\alpha}(t(u-1))^{w_\alpha(-t)}, \end{equation}
equals $w_\alpha(tu) \cdot w_\alpha(-t)$. This allows us to set $\{t, u \}'_\alpha := f(\alpha, t, u) f(\alpha, 1, u)^{-1}.$
Now, obviously $\mu_I\{ t, u \}'_\alpha = w_\alpha(tu) w_\alpha(-t) w_\alpha(1) w_\alpha(-u) = \{t, u\}_\alpha,$ as claimed.
\begin{rem}
 Our relative Steinberg symbols differs from the one defined in~\cite{Ste73} in that
  it takes values in $\St(\Phi, R, I)$ rather than the group $\St(\Phi, R)$. %Elaborate on this...
\end{rem}

\begin{comment}
 \{ t, u \}'_\alpha = h_{\alpha}(u) ^ {h_{?}(t)} =  y_\alpha(t(u-1)) \cdot y_{-\alpha}\left(-t^{-1}(u^{-1}-1)\right)^{x_\alpha(-t)} \cdot y_\alpha\left(t(u-1)\right)^{w_\alpha(-t)} \cdot \\
\cdot y_\alpha(u-1) \cdot y_{-\alpha}\left(-(u^{-1}-1)\right)^{x_\alpha(-1)} \cdot y_\alpha(u-1)^{w_\alpha(-1)}.
\end{comment}

@@

\begin{rem} Despite the fact that formula~\eqref{eq:sigmadef} determines the maps $\sigma(\varpi)$, $\sigma(-\varpi)$ uniquely, 
 we cannot a priori be sure that~\eqref{eq:sigmadef} actually {\it defines} them, i.\,e. gives well-defined maps between $N(\pm\varpi)$.
This is caused by the fact there is no direct description of the set of relations which hold in $N(\pm\varpi)$ between the generators $\XX(\pm\varpi)$, 
 so we cannot verify that $\sigma(\pm\varpi)$ preserve these relations. \end{rem}

@@

Denote by $V_0$ the set $\St(\Phi, A[t]) \times \StH(A[t, t^{-1}]) \times \St(\Phi, A[t, t^{-1}], m[t, t^{-1}])$. 
We define an equivalence relation $\sim$ on $V_0$ as the minimal equivalence relation generated by the following 3 elementary transformations:
\begin{enumerate}
 \item $(a \cdot u, b, \beta) \sim (a, j_+^*(u) \cdot b, \beta)$, where $u$ is an element $\UU(\Phi^+, A[t])$;
 \item $(a \cdot \ext{m[t]}(\gamma), b, \beta) \sim (a, b, j^*_+(\gamma)^b \cdot \beta)$, where $\gamma$ is an element of $\St(\Phi, A[t], m[t])$;
 \item $(a, b \cdot u, \beta) \sim (a, b, u\cdot \beta)$, where $u$ is an element of $\UU(\Phi^+, m[t, t^{-1}])$;
 \item $(a, b \cdot \{ t, a \}_\alpha, \beta) \sim (a, b, \{ t, a \}'_\alpha \cdot \beta) $, where $a \in (1+m)^*$
  and $\{t, a\}'_\alpha$ is a relative Steinberg symbol discussed in~\cref{sec:relative-steinberg}
\end{enumerate}
